Four Models for Severe Weather Deaths

### 2017-05-11 paper.Rnw
Edited for length, fixed problem with figures

\setlength{\marginparwidth}{2cm}
\begin{document}
\SweaveOpts{concordance=TRUE}
<<imports, cache=FALSE, echo=FALSE>>= 
read_chunk('zinb_sampler.R')
options(scipen=4, digits=3)
library('truncnorm')
@
<<dataset, echo=FALSE, cache=FALSE>>= 
load('small_storms.RData')
storm_details <- small_storms
@
\begin{center}
	\vspace{0.1cm}
	\textsc{\LARGE MATH 640 Final Project} \\[0.1cm]
	Jason Michaels (jam521), Niko Paulson (ndp32), Miranda Seitz-McLeese (mgs85) 
\end{center}
\section{Introduction}
\label{s:intro}
This analysis will be done on a data set of a variety of measurements about severe weather in the United States. 
The data set contains a variety of measures from severe weather events in the United States from 1996-2016. It was taken from the NOAA website \cite{dataset}.
For this project we focused on the deaths directly attributable to the event. In the interest of time we only modeled two types of severe weather: Tornadoes and Flash Floods.
Understanding how and at what rate severe weather events become lethal in the United States has tremendous public health impacts.
In this paper we compare four possible models for the deaths: The traditional Poisson and negative binomial distributions, as well as the zero inflated variant of each.

@@

Fortunately, the vast majority of severe weather events in the United States involve no deaths, therefore we wanted to also account for the possibility of structural zeros, therefore we also considered zero inflated variants. 
These distributions are created by returning $0$ with probability $\sigma$ and sampling from the original distribution with probability $(1-\sigma)$.

We will derive and fit a model for each of the four distributions and see if there is a difference in our results and evaluate to determine which model best fits the data.

@@

For each variable of interest, 20,000 samples were taken. As convergence was not immediately achieved, the first 10,000 samples were discarded as a burn-in. The results are summarized in Table~\ref{t:rZIP}.

\begin{table}
    \centering
    \caption{Results of taking 20,000 samples from the posterior distributions of $\lambda$ and p, compared between the two event types. The second column relays the mean of the sample for $\lambda$, with the 95 percent credible interval in parentheses. The third does the same for p. All means and credible intervals are taken after discarding the first 10,000 samples as a burn-in.}
    \label{t:rZIP}
    \begin{tabular}{| l | l | l |}
    \hline
    Event Type & $\lambda$ & p  \\ \hline
    Tornado & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
    Flash Flood & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
    \end{tabular}
\end{table}

\includegraphics{figure/ZIP_Flood_Conv.pdf}

\noindent Figure 1. Cumulative mean plots for the posterior samples of the two parameters from the flood data. Plots for the tornado data are not pictured, but behave similarly.

\includegraphics{figure/ZIP_Density.pdf}

\noindent Figure 2. Posterior density plots for flood parameters (in blue), and tornado parameters (black).

@@

This model suffered severely from the curse of dimensionality. It was very slow to converge, and the very small variance (necessary to tune the acceptance rate) led to high autocorrelation, meaning agressive thinning was necessary. Autocorrelation and running mean plots the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}. 
<<floodsample, cache=TRUE, cache.vars='flood.zinb', echo=FALSE, dependson='zinbparams'>>=
flood.zinb <- zinb.sampler(storm_details, "Flash Flood", samplesize, 
                           r.var=r.var, p.var=p.var, sigma.var=sigma.var, 
                           burnin=burnin, thinning=thin)
@
<<tornadosample, cache=TRUE, cache.vars='tornado.zinb', echo=FALSE, dependson='zinbparams'>>=
tornado.zinb <- zinb.sampler(storm_details, "Tornado", samplesize, 
                             r.var=r.var, p.var=p.var, sigma.var=sigma.var, 
                             burnin=burnin, thinning=thin)
@
In the end $\Sexpr{samplesize}$ samples were drawn originally for each model, with a burn in of $\Sexpr{burnin}$ and one of every $\Sexpr{thin}$ samples was retained, resulting in a final sample size of $\Sexpr{length(flood.zinb$sigma)}$. The flood model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$ and the tornado model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$. The resulting densities are pictured in Figure~\ref{f:zinbdensity}, and Table~\ref{t:ZINBparams}, though due to autocorrelation and questions of convergence the results should viewed with caution.
<<evaluation, echo=FALSE, cache=FALSE>>=
@
<<model-eval, echo=FALSE, cache=TRUE, dependson=c('floodsample', 'tornadosample')>>=
flood.dic.zinb <- dic(storm_details, "Flash Flood", flood.zinb$sigma, 
                      flood.zinb$r, flood.zinb$p)
tornado.dic.zinb <- dic(storm_details, "Tornado", tornado.zinb$sigma, 
                        tornado.zinb$r, tornado.zinb$p)
#flood.wiac.zinb <- wiac(storm_details, "Flash Flood", flood.zinb$sigma, 
#                   flood.zinb$r, flood.zinb$p)
#tornado.wiac.zinb <- wiac(storm_details, "Tornado", tornado.zinb$sigma, 
#                          tornado.zinb$r, tornado.zinb$p)
@

The DIC for floods and tornadoes respectively were $\Sexpr{as.integer(flood.dic.zinb)}$ and $\Sexpr{as.integer(tornado.dic.zinb)}$. I was unable to calculate WIACs because it was too computationally expensive.
\begin{table}
\centering
\caption{The following table shows the results of the flash flood ZINB model and the tornado ZINB model. It shows the mean and the median for each parameter as well as the 95\% credible interval.}
\label{t:ZINBparams}
\begin{tabular}{lcccccc}
\toprule
&\multicolumn{3}{c}{Flash Flood}&\multicolumn{3}{c}{Tornado}\\
\cmidrule(r){2-4}\cmidrule(l){5-7}
Parameter & Mean Value &Median Value & (95\% CI)& Mean Value &Median Value & (95\% CI)\\
\midrule
$\sigma$ & \Sexpr{mean(flood.zinb$sigma)} &
  \Sexpr{median(flood.zinb$sigma)}&
  (\Sexpr{quantile(flood.zinb$sigma, probs=c(.025))[1]}, 
    \Sexpr{quantile(flood.zinb$sigma, probs=c(.975))[1]})&
  \Sexpr{mean(tornado.zinb$sigma)} &
  \Sexpr{median(tornado.zinb$sigma)}&
  (\Sexpr{quantile(tornado.zinb$sigma, probs=c(.025))[1]}, 
    \Sexpr{quantile(tornado.zinb$sigma, probs=c(.975))[1]})\\
$p$ &  \Sexpr{mean(flood.zinb$p)} &
  \Sexpr{median(flood.zinb$p)}&
  (\Sexpr{quantile(flood.zinb$p, probs=c(.025))[1]}, 
    \Sexpr{quantile(flood.zinb$p, probs=c(.975))[1]})&
  \Sexpr{mean(tornado.zinb$p)} &
  \Sexpr{median(tornado.zinb$p)}&
  (\Sexpr{quantile(tornado.zinb$p, probs=c(.025))[1]}, 
    \Sexpr{quantile(tornado.zinb$p, probs=c(.975))[1]})\\
$r$ &  \Sexpr{mean(flood.zinb$r)} &
  \Sexpr{median(flood.zinb$r)}&
  (\Sexpr{quantile(flood.zinb$r, probs=c(.025))[1]}, 
    \Sexpr{quantile(flood.zinb$r, probs=c(.975))[1]})&
  \Sexpr{mean(tornado.zinb$r)} &
  \Sexpr{median(tornado.zinb$r)}&
  (\Sexpr{quantile(tornado.zinb$r, probs=c(.025))[1]}, 
    \Sexpr{quantile(tornado.zinb$r, probs=c(.975))[1]})\\
\bottomrule
\end{tabular}
\end{table}
\begin{figure}
	\centering
<<floodconvergence, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=3.5>>=
B<-length(flood.zinb$sigma)
Bt<-length(tornado.zinb$sigma)
par(mfrow=c(2, 3))
par(mgp=c(1.25,.5,0))
par(mar=c(3.1,4.1,2.5,.1))
plot(cumsum(flood.zinb$sigma)/1:B, 
     main = expression(paste('Cum. Mean (Flood) ',sigma)),
     ylab = 'Cumulative Mean',
     type='l')
plot(cumsum(flood.zinb$r)/1:B, 
     main = expression(paste('Cum. Mean (Flood) ',r)),
     ylab = 'Cumulative Mean',
     type='l')
plot(cumsum(flood.zinb$p)/1:B, 
     type='l', 
     main = expression(paste('Cum. Mean (Flood) ',p)),
     ylab = 'Cumulative Mean')
acf(flood.zinb$sigma, 
    main = expression(paste('ACF (Flood) ',sigma)))
acf(flood.zinb$r,
    main = expression(paste('ACF (Flood) ',r)))
acf(flood.zinb$p,
    main = expression(paste('ACF (Flood) ',p)))
# plot(cumsum(tornado.zinb$sigma)/1:Bt, 
#      main = expression(paste('Cum. Mean (Tornado) ',sigma)),
#      ylab = 'Cumulative Mean',
#      type='l')
# plot(cumsum(tornado.zinb$r)/1:Bt, 
#      main = expression(paste('Cum. Mean (Tornado) ',r)),
#      ylab = 'Cumulative Mean',
#      type='l')
# plot(cumsum(tornado.zinb$p)/1:Bt, 
#      type='l', 
#      main = expression(paste('Cum. Mean (Tornado) ',p)),
#      ylab = 'Cumulative Mean')
# acf(tornado.zinb$sigma, 
#     main = expression(paste('ACF (Tornado) ',sigma)))
# acf(tornado.zinb$r,
#     main = expression(paste('ACF (Tornado) ',r)))
# acf(tornado.zinb$p,
#     main = expression(paste('ACF (Tornado) ',p)))
@
	\caption{Autocorrelation plots, and cumulative mean plots for each of the three parameters in the ZINB model fit on flash flood data. The tornado plots are not pictured, but are similar in character.}
	\label{f:zinbconvergence}
\end{figure}

\begin{figure}
	\centering
<<zinbdensity, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=3>>=
par(mfrow=c(1,3))
double.density.plot(flood.zinb$sigma, tornado.zinb$sigma,
                    expression(paste(sigma, ' Posterior Density ')), 
                    expression(sigma), 
                    expression(p*group("(",paste(sigma, "|", X),")")),
                    col=c('blue', 'black'))
double.density.plot(flood.zinb$p, tornado.zinb$p,
                    expression(paste(p, ' Posterior Density ')), 
                    expression(p), 
                    expression(p*group("(",paste(p, "|", X),")")),
                    col=c('blue', 'black'))
double.density.plot(flood.zinb$r, tornado.zinb$r,
                    expression(paste(r, ' Posterior Density ')), 
                    expression(r), 
                    expression(p*group("(",paste(r, "|", X),")")),
                    col=c('blue', 'black'))
@
	\caption{Posterior density curves for each of the three variables. The flood posterior is shown in blue and the tornado posterior is shown in black.}
	\label{f:zinbdensity}
\end{figure} 
\section{Discussion}
\label{s:discussion}
The goal of this analysis was to determine which of the four models was the best fit for the data. In order to make this determination the DIC and WAIC was calculated for each model on both tornados and flash floods. The results are summarized in Table~\ref{t:evalresults}.

\begin{table}
\centering
\caption{The following table shows the resulting DIC and WAIC for each of the four models on both the tornado data and the flash flood data. In the case of the zero inflated negative binomial model, WAIC was too computationally expensive, so it was not calculated.}
\label{t:evalresults}
\begin{tabular}{lcccc}
\toprule
&\multicolumn{2}{c}{Flash Flood}&\multicolumn{2}{c}{Tornado}\\
\cmidrule(r){2-3}\cmidrule(l){4-5}
Model & DIC &WAIC & DIC &WAIC\\
\midrule
Poisson & P FLOOD DIC& P FLOOD WAIC &P TORNADO DIC&P TORNADO WAIC\\
Negative Binomial & NB FLOOD DIC& NB FLOOD WAIC &NB TORNADO DIC&NB TORNADO WAIC\\
ZIP & 3845.915& ZIP FLOOD WAIC &4759.998 &ZIP TORNADO WAIC\\
ZINB & \Sexpr{flood.dic.zinb}& -- &\Sexpr{tornado.dic.zinb}&--\\
\bottomrule
\end{tabular}
\end{table}

\begin{thebibliography}{2}
  \bibitem{dataset} NOAA's Severe Weather Data Inventory, 
    \url{https://www1.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/}. 
    Accessed April 2017.
  \bibitem{ZIP} Bayarri, M., Berger, J., Datta, G. (2008). Objective testing of Poisson versus inflated Poisson models. IMS 	Collections, 3, 105-121. 
\end{thebibliography}
\clearpage
\appendix
\section{Code}
\label{a:code}
This appendix includes the code used to implement the models.
\subsection{Zero Inflated Negative Binomial code}
\label{a:cZINB}
<<zinbsampler-source, echo=TRUE, ref.label='zinbsampler'>>=
@
\subsection{Zero Inflated Poisson code}
<<eval = FALSE>>=
library(dplyr)
tornado <- data %>% filter(EVENT_TYPE == "Tornado") %>% select(DEATHS_DIRECT)
flood <- data %>% filter(EVENT_TYPE == "Flash Flood") %>% select(DEATHS_DIRECT)

>>>>

\setlength{\marginparwidth}{2cm}
\begin{document}
<<imports, cache=FALSE, echo=FALSE>>= 
read_chunk('zinb_sampler.R')
options(scipen=4, digits=3)
library('truncnorm')
@
<<dataset, echo=FALSE, cache=FALSE>>= 
load('small_storms.RData')
storm_details <- small_storms
@
\begin{center}
	\vspace{0.1cm}
	\textsc{\LARGE MATH 640 Final Project} \\[0.1cm]
	Jason Michaels (jam521), Niko Paulson (ndp32), Miranda Seitz-McLeese (mgs85) 
\end{center}
\section{Introduction}
\label{s:intro}
This analysis will be done on a data set of a variety of measurements about severe weather in the United States. 
The data set contains a variety of measures from severe weather events in the United States from 1996-2016. It was taken from the NOAA website \cite{dataset}.
For this project we focused on the deaths directly attributable to the event. In the interest of time we only modeled two types of severe weather: Tornadoes and Flash Floods.
Understanding how and at what rate severe weather events become lethal in the United States has tremendous public health impacts.
In this paper we compare four possible models for the deaths: The traditional Poisson and negative binomial distributions, as well as the zero inflated variant of each.

@@

Fortunately, the vast majority of severe weather events in the United States involve no deaths, therefore we wanted to also account for the possibility of structural zeros, therefore we also considered zero inflated variants. 
These distributions are created by returning $0$ with probability $\sigma$ and sampling from the original distribution with probability $(1-\sigma)$. We will derive and fit a model for each of the four distributions and see if there is a difference in our results and evaluate to determine which model best fits the data.

@@

For each variable of interest, 20,000 samples were taken. As convergence was not immediately achieved, the first 10,000 samples were discarded as a burn-in. The results are summarized in Table~\ref{t:rZIP}. Convergance plots can be found in Figrue~\ref{f:zipconv}, and density plots can be found in Figure~\ref{f:zipdensity}.

\begin{table}
    \centering
    \caption{The posterior distributions of $\lambda$ and p, for both event types. The second column relays the mean of the sample for $\lambda$, with the 95 percent credible interval in parentheses. The third does the same for p.}
    \label{t:rZIP}
    \begin{tabular}{| l | l | l |}
    \hline
    Event Type & $\lambda$ & p  \\ \hline
    Tornado & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
    Flash Flood & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
    \end{tabular}
\end{table}

\begin{figure}[p]
\centering
\includegraphics[width=.65\textwidth]{figure/ZIP_Flood_Conv.pdf}
\caption{Cumulative mean plots for the posterior samples (ZIP model) of the two parameters from the flood data. Plots for the tornado data are not pictured, but behave similarly.}
\label{f:zipconv}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[width=.65\textwidth]{figure/ZIP_Density.pdf}
\caption{Posterior density plots (ZIP model) for flood parameters (in blue), and tornado parameters (black). The solid line represents the mean and the dashed lines indicate the 95\% confidence interval.}
\label{f:zipdensity}
\end{figure}

@@

This model suffered severely from the curse of dimensionality. It was very slow to converge, and the very small variance needed for a reasonable acceptance led to high autocorrelation, meaning agressive thinning was necessary. Running mean plots the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}. 
<<floodsample, cache=TRUE, cache.vars='flood.zinb', echo=FALSE, dependson='zinbparams'>>=
flood.zinb <- zinb.sampler(storm_details, "Flash Flood", samplesize, 
                           r.var=r.var, p.var=p.var, sigma.var=sigma.var, 
                           burnin=burnin, thinning=thin)
@
<<tornadosample, cache=TRUE, cache.vars='tornado.zinb', echo=FALSE, dependson='zinbparams'>>=
tornado.zinb <- zinb.sampler(storm_details, "Tornado", samplesize, 
                             r.var=r.var, p.var=p.var, sigma.var=sigma.var, 
                             burnin=burnin, thinning=thin)
@
In the end $\Sexpr{samplesize}$ samples were drawn for each model, with a burn in of $\Sexpr{burnin}$ and one of every $\Sexpr{thin}$ samples was retained, resulting in a final sample size of $\Sexpr{length(flood.zinb$sigma)}$. The flood model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$ and the tornado model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$. The resulting densities are pictured in Figure~\ref{f:zinbdensity}, and Table~\ref{t:ZINBparams}, though due to questions of convergence the results should viewed with caution.
<<evaluation, echo=FALSE, cache=FALSE>>=
@
<<model-eval, echo=FALSE, cache=TRUE, dependson=c('floodsample', 'tornadosample')>>=
flood.dic.zinb <- dic(storm_details, "Flash Flood", flood.zinb$sigma, 
                      flood.zinb$r, flood.zinb$p)
tornado.dic.zinb <- dic(storm_details, "Tornado", tornado.zinb$sigma, 
                        tornado.zinb$r, tornado.zinb$p)
#flood.wiac.zinb <- wiac(storm_details, "Flash Flood", flood.zinb$sigma, 
#                   flood.zinb$r, flood.zinb$p)
#tornado.wiac.zinb <- wiac(storm_details, "Tornado", tornado.zinb$sigma, 
#                          tornado.zinb$r, tornado.zinb$p)
@

The DIC for floods and tornadoes respectively were $\Sexpr{as.integer(flood.dic.zinb)}$ and $\Sexpr{as.integer(tornado.dic.zinb)}$.
\begin{table}
\centering
\caption{The following table shows the results of the flash flood ZINB model and the tornado ZINB model. It shows the mean and the median for each parameter as well as the 95\% credible interval.}
\label{t:ZINBparams}
\begin{tabular}{lcccccc}
\toprule
&\multicolumn{3}{c}{Flash Flood}&\multicolumn{3}{c}{Tornado}\\
\cmidrule(r){2-4}\cmidrule(l){5-7}
Parameter & Mean Value &Median Value & (95\% CI)& Mean Value &Median Value & (95\% CI)\\
\midrule
$\sigma$ & \Sexpr{mean(flood.zinb$sigma)} &
  \Sexpr{median(flood.zinb$sigma)}&
  (\Sexpr{quantile(flood.zinb$sigma, probs=c(.025))[1]}, 
    \Sexpr{quantile(flood.zinb$sigma, probs=c(.975))[1]})&
  \Sexpr{mean(tornado.zinb$sigma)} &
  \Sexpr{median(tornado.zinb$sigma)}&
  (\Sexpr{quantile(tornado.zinb$sigma, probs=c(.025))[1]}, 
    \Sexpr{quantile(tornado.zinb$sigma, probs=c(.975))[1]})\\
$p$ &  \Sexpr{mean(flood.zinb$p)} &
  \Sexpr{median(flood.zinb$p)}&
  (\Sexpr{quantile(flood.zinb$p, probs=c(.025))[1]}, 
    \Sexpr{quantile(flood.zinb$p, probs=c(.975))[1]})&
  \Sexpr{mean(tornado.zinb$p)} &
  \Sexpr{median(tornado.zinb$p)}&
  (\Sexpr{quantile(tornado.zinb$p, probs=c(.025))[1]}, 
    \Sexpr{quantile(tornado.zinb$p, probs=c(.975))[1]})\\
$r$ &  \Sexpr{mean(flood.zinb$r)} &
  \Sexpr{median(flood.zinb$r)}&
  (\Sexpr{quantile(flood.zinb$r, probs=c(.025))[1]}, 
    \Sexpr{quantile(flood.zinb$r, probs=c(.975))[1]})&
  \Sexpr{mean(tornado.zinb$r)} &
  \Sexpr{median(tornado.zinb$r)}&
  (\Sexpr{quantile(tornado.zinb$r, probs=c(.025))[1]}, 
    \Sexpr{quantile(tornado.zinb$r, probs=c(.975))[1]})\\
\bottomrule
\end{tabular}
\end{table}
\begin{figure}[p]
	\centering
<<floodconvergence, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=1.75>>=
B<-length(flood.zinb$sigma)
Bt<-length(tornado.zinb$sigma)
par(mfrow=c(1, 3))
par(mgp=c(1.25,.5,0))
par(mar=c(3.1,4.1,2.5,.1))
plot(cumsum(flood.zinb$sigma)/1:B, 
     main = expression(paste('Cum. Mean (Flood) ',sigma)),
     ylab = 'Cumulative Mean',
     type='l')
plot(cumsum(flood.zinb$r)/1:B, 
     main = expression(paste('Cum. Mean (Flood) ',r)),
     ylab = 'Cumulative Mean',
     type='l')
plot(cumsum(flood.zinb$p)/1:B, 
     type='l', 
     main = expression(paste('Cum. Mean (Flood) ',p)),
     ylab = 'Cumulative Mean')
# acf(flood.zinb$sigma, 
#     main = expression(paste('ACF (Flood) ',sigma)))
# acf(flood.zinb$r,
#     main = expression(paste('ACF (Flood) ',r)))
# acf(flood.zinb$p,
#     main = expression(paste('ACF (Flood) ',p)))
# plot(cumsum(tornado.zinb$sigma)/1:Bt, 
#      main = expression(paste('Cum. Mean (Tornado) ',sigma)),
#      ylab = 'Cumulative Mean',
#      type='l')
# plot(cumsum(tornado.zinb$r)/1:Bt, 
#      main = expression(paste('Cum. Mean (Tornado) ',r)),
#      ylab = 'Cumulative Mean',
#      type='l')
# plot(cumsum(tornado.zinb$p)/1:Bt, 
#      type='l', 
#      main = expression(paste('Cum. Mean (Tornado) ',p)),
#      ylab = 'Cumulative Mean')
# acf(tornado.zinb$sigma, 
#     main = expression(paste('ACF (Tornado) ',sigma)))
# acf(tornado.zinb$r,
#     main = expression(paste('ACF (Tornado) ',r)))
# acf(tornado.zinb$p,
#     main = expression(paste('ACF (Tornado) ',p)))
@
	\caption{Cumulative mean plots for each of the three parameters in the ZINB model fit on flash flood data. The tornado plots are not pictured, but are similar in character.}
	\label{f:zinbconvergence}
\end{figure}

\begin{figure}[p]
	\centering
<<zinbdensity, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=2.5, fig.width=6>>=
par(mfrow=c(1,3))
double.density.plot(flood.zinb$sigma, tornado.zinb$sigma,
                    expression(paste(sigma, ' Posterior Density ')), 
                    expression(sigma), 
                    expression(p*group("(",paste(sigma, "|", X),")")),
                    col=c('blue', 'black'))
double.density.plot(flood.zinb$p, tornado.zinb$p,
                    expression(paste(p, ' Posterior Density ')), 
                    expression(p), 
                    expression(p*group("(",paste(p, "|", X),")")),
                    col=c('blue', 'black'))
double.density.plot(flood.zinb$r, tornado.zinb$r,
                    expression(paste(r, ' Posterior Density ')), 
                    expression(r), 
                    expression(p*group("(",paste(r, "|", X),")")),
                    col=c('blue', 'black'))
@
	\caption{Posterior density curves for each of the three variables. The flood posterior is shown in blue and the tornado posterior is shown in black.}
	\label{f:zinbdensity}
\end{figure} 
\section{Discussion}
\label{s:discussion}
The goal of this analysis was to determine which of the four models was the best fit for the data. In order to make this determination the DIC and WAIC was calculated for each model on both tornados and flash floods. The results are summarized in Table~\ref{t:evalresults}.

\begin{table}
\centering
\caption{The following table shows the resulting DIC for each of the four models on both the tornado data and the flash flood data.}
\label{t:evalresults}
\begin{tabular}{lcc}
\toprule
Model & DIC Flash Flood & DIC Tornado\\
\midrule
Poisson & P FLOOD DIC& P TORNADO DIC\\
Negative Binomial & NB FLOOD DIC&NB TORNADO DIC\\
ZIP & 10454.860&8727.647 \\
ZINB & \Sexpr{flood.dic.zinb}&\Sexpr{tornado.dic.zinb}\\
\bottomrule
\end{tabular}
\end{table}
\clearpage
\begin{thebibliography}{2}
  \bibitem{dataset} NOAA's Severe Weather Data Inventory, 
    \url{https://www1.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/}. 
    Accessed April 2017.
  \bibitem{ZIP} Bayarri, M., Berger, J., Datta, G. (2008). Objective testing of Poisson versus inflated Poisson models. IMS 	Collections, 3, 105-121. 
\end{thebibliography}

\appendix
\section{Code}
\label{a:code}
This appendix includes the code used to implement the models.
\subsection{Zero Inflated Negative Binomial code}
\label{a:cZINB}
<<zinbsampler-source, echo=TRUE, ref.label='zinbsampler'>>=
@
\subsection{Zero Inflated Poisson code}
<<eval = FALSE>>=
library(dplyr)
tornado <- data %>% filter(EVENT_TYPE == "Tornado") %>% select(DEATHS_DIRECT)
flood <- data %>% filter(EVENT_TYPE == "Flash Flood") %>% select(DEATHS_DIRECT)
